---
title: "SuCAR_preprocessing"
author: "Maria Meier"
date: "`r Sys.Date()`"
output: html_document
---

**Abstract**

The hypothalamic-pituitary-adrenal (HPA) axis is involved in vital physiological processes, and its regulation has been associated with mental health impairments. The cortisol awakening response (CAR) is considered a naturally occurring reactivity marker of the HPA axis and describes a rapid increase in cortisol levels in response to awakening. While cortisol surges are proposed to mobilize energy, e.g. to support adaptation to acute stress or to prepare the organism for the upcoming day, it has also been shown that glucose intake modulates cortisol reactivity in the context of psychosocial stress: Glucose increases net cortisol output, overall amplifying cortisol stress reactivity. The question of whether this amplificatory effect also occurs in the context of the CAR has not been tested experimentally to date. This project therefore investigates whether glucose intake after awakening increases the CAR. In a placebo-controlled, ambulatory within-subject study, we experimentally manipulate blood glucose concentration after waking up by administering standardized drinks (glucose vs. sweetener). Concurrently, participants will collect four salivary samples to capture the CAR and use continuous glucose monitoring sensors to capture the dynamics of glucose metabolism in high resolution. By doing so, we will be able to link changes in glucose levels after awakening to the CAR. As both cortisol regulation as well as insulin sensitivity varies across the menstrual cycle, we consider possible modulating effects of sex and menstrual phase in our analyses. The results of this project inform on potential links between glucose metabolism and naturally occurring cortisol reactivity, which may deepen our understanding of how metabolic factors may impact endocrine regulation and the risk for mental health impairments.

**Preregistration**
The hypotheses and analysis plan was preregistered on Open Science Framework on .....

# Merging and preprocessing the data

## Setup

```{r packaged}
library(tidyverse)
library(psych)
library(openxlsx)
library(reshape2)
library(readxl)
library(stringr)
library(dplyr)
library(emmeans)
library(ggplot2)
library(nlme)
```

```{r}
project <- "SuCAR"
wd <- "~/Nextcloud/AG Pruessner/Projekte/SuCAR/5. Data and analysis/"
```

## Load data

### Tagebuch

```{r}
# protocol data: code and vp number
data.protocol <- read.csv("/Users/maria/Nextcloud/AG Pruessner/Projekte/SuCAR/5. Data and analysis/1 Rawdata/Tagebuch/SuCAR_Messtagebuch_20251007.csv")
data.protocol <- data.protocol[-1, ]
names(data.protocol)[names(data.protocol) == "number"] <- "vp_nr"
# correct duplicated code
data.protocol$code <- toupper(data.protocol$code)
```

### Salivary cortisol

```{r}
# define working directory
work.dir <- "~/Nextcloud/AG Pruessner/Projekte/SuCAR/5. Data and analysis/2 Cleaned Data/"
setwd(work.dir)

# create subdirectory for cortisol in 2 Cleaned Data
dir.create("Cortisol")
work.dir <- paste(work.dir, "Cortisol/", sep = "")
setwd(work.dir)

# read in excel you received from wetlab
cortisol_raw <- read_excel("~/Nextcloud/AG Pruessner/Projekte/SuCAR/5. Data and analysis/1 Rawdata/Cortisol/SuCAR_Cort_results_20251007.xlsx")

# read out inter and intra CV
interCV <- round(mean(as.numeric(names(cortisol_raw)[6])), 2)

intraCV_col <- cortisol_raw[, 8]
names(intraCV_col) <- "intra"
intraCV_col$intra[intraCV_col$intra == "intra %CV"] <- NA
intraCV_col$intra <- as.numeric(intraCV_col$intra)
intraCV <- round(mean(intraCV_col$intra, na.rm = TRUE), 2)

CV_data <- data.frame(matrix(data = NA, nrow = 1, ncol = 2)); names(CV_data) <- c("interCV", "intraCV")
CV_data$interCV <- interCV; CV_data$intraCV <- intraCV 

# save inter and intra CV means in file
write.csv(x = CV_data, 
          file = paste(project, "_cort_assay_CVs_", Sys.Date(), ".csv", sep = ""))
rm(interCV); rm(intraCV); rm(CV_data)

# write cortisol data into dataframe (long) - csv and RData
subset.cort.long <- cortisol_raw[, c(2, 3, 7)]

colnames(subset.cort.long) <- c("vp_nr", "variable", "cortisol")
subset.cort.long <- subset.cort.long[-which(subset.cort.long$vp_nr %in% c("Label", NA)),]

subset.cort.long$vp_nr <- ifelse(as.numeric(subset.cort.long$vp_nr) < 10, paste0("VP0", subset.cort.long$vp_nr), paste0("VP", subset.cort.long$vp_nr))

subset.cort.long$vp_nr <- as.factor(subset.cort.long$vp_nr)
subset.cort.long$cortisol <- as.numeric(subset.cort.long$cortisol)
subset.cort.long$day <- substring(subset.cort.long$variable, 1, 1)
subset.cort.long$day <- as.factor(subset.cort.long$day)
subset.cort.long$variable <- substring(subset.cort.long$variable, 2)
subset.cort.long$variable <- paste("Cort_", subset.cort.long$variable, sep = "")
colnames(subset.cort.long) <- c("vp_nr", "variable", "cortisol", "day")

# insert blinding
blind <- read.csv("~/Nextcloud/AG Pruessner/Projekte/SuCAR/4. Testing/1 Testungen/Randomisierung/SuCAR_random_allocation.csv")
subjects <- unique(subset.cort.long$vp_nr)
names(blind)[names(blind) == "subject_number"] <- "vp_nr"
blind$vp_nr <- ifelse(as.numeric(blind$vp_nr) < 10, paste0("VP0", blind$vp_nr), paste0("VP", blind$vp_nr))
blind <- blind[blind$vp_nr %in% subjects, ]
blind$A <- sapply(strsplit(blind$conditions, ","), "[[", 1)
blind$B <- str_trim(sapply(strsplit(blind$conditions, ","), "[[", 2))
blind$C <- str_trim(sapply(strsplit(blind$conditions, ","), "[[", 3))
blind$D <- str_trim(sapply(strsplit(blind$conditions, ","), "[[", 4))

blind.long <- melt(blind[, c("vp_nr", "A", "B", "C", "D")], 
                   id.vars = "vp_nr")
names(blind.long)[names(blind.long) == "variable"] <- "day"
names(blind.long)[names(blind.long) == "value"] <- "condition"

subset.cort.long <- merge(subset.cort.long, blind.long, 
                          by = c("vp_nr", "day"))
rm(blind); rm(blind.long); rm(cortisol_raw)

# insert timing
logs <- read.csv("/Users/maria/Nextcloud/AG Pruessner/Projekte/SuCAR/5. Data and analysis/2 Cleaned Data/CARWatch Log Files/SuCAR_logs_combined_05-09-2025.csv")
names(logs)[names(logs) == "Participant.ID"] <- "vp_nr" 
logs$vp_nr <- toupper(logs$vp_nr)
# VP 20 scanned teh first day twice - correct
logs <- logs[-240,]
logs$sample_label[logs$vp_nr == "VP20" & logs$date_D1 == "2025-06-26"] <- c("B1_D1", "B2_D1", "B3_D1", "B4_D1")
#readr::parse_number(logs$Participant.ID)

logs$day <- substr(logs$sample_label, start = 1, stop = 1)
logs$variable <- paste0("Cort_", logs$sample_barcode %% 10)

subset.cort.long <- merge(subset.cort.long, 
                          logs[, c("vp_nr", "variable", "date_D1", "awakening_time_D1_app", 
                                   "sampling_time", "day", "variable")], 
                          by = c("vp_nr", "day", "variable"),
                          all = TRUE)

subset.cort.long <- subset.cort.long[, c("vp_nr", "day", "variable", "cortisol", "condition", "date_D1", "awakening_time_D1_app", "sampling_time")]

subset.cort.long$opt_lag[subset.cort.long$variable == "Cort_1"] <- 0 
subset.cort.long$opt_lag[subset.cort.long$variable == "Cort_2"] <- 30
subset.cort.long$opt_lag[subset.cort.long$variable == "Cort_3"] <- 45
subset.cort.long$opt_lag[subset.cort.long$variable == "Cort_4"] <- 60

# Deal with missing or incorrect logs

# missing logs from app
missinglogs <- data.frame(cbind(as.character(subset.cort.long$vp_nr[is.na(subset.cort.long$sampling_time)]), as.character(subset.cort.long$day[is.na(subset.cort.long$sampling_time)])))
names(missinglogs) <- c("vp_nr", "day")
missinglogs <- unique(missinglogs)

# manual logs from protocol
data.protocol$sampling_time_manual[data.protocol$sampling_time_manual == ""] <- 0
data.protocol$sampling_time_manual[is.na(data.protocol$sampling_time_manual)] <- 0
manuallogs <- data.frame(cbind(as.character(data.protocol$vp_nr[data.protocol$sampling_time_manual == "1"]), as.character(data.protocol$day[data.protocol$sampling_time_manual == "1"])))
names(manuallogs) <- c("vp_nr", "day")

subset.cort.long$awakening_time_D1_app <- as.POSIXct(paste(subset.cort.long$date_D1, subset.cort.long$awakening_time_D1_app, sep = " "), format = "%Y-%m-%d %H:%M:%S")
subset.cort.long$sampling_time <- as.POSIXct(paste(subset.cort.long$date_D1, subset.cort.long$sampling_time, sep = " "), format = "%Y-%m-%d %H:%M:%S")
subset.cort.long$lag <- as.numeric((subset.cort.long$sampling_time - subset.cort.long$awakening_time_D1_app)/60)
subset.cort.long$lag_min <- subset.cort.long$lag - subset.cort.long$opt_lag

# for those who provided info, insert manual logs to log file
## VP06, day B
subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP06" & subset.cort.long$day == "B")] <- c(data.protocol$sampling_time_1[which(data.protocol$vp_nr == "VP06" & data.protocol$day == "B")], data.protocol$sampling_time_2[which(data.protocol$vp_nr == "VP06" & data.protocol$day == "B")], data.protocol$sampling_time_3[which(data.protocol$vp_nr == "VP06" & data.protocol$day == "B")], data.protocol$sampling_time_4[which(data.protocol$vp_nr == "VP06" & data.protocol$day == "B")])
subset.cort.long$date_D1[which(subset.cort.long$vp_nr == "VP06" & subset.cort.long$day == "B")] <- "2025-05-29"

## VP10, day A
subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP10" & subset.cort.long$day == "A")] <- c(data.protocol$sampling_time_1[which(data.protocol$vp_nr == "VP10" & data.protocol$day == "A")], data.protocol$sampling_time_2[which(data.protocol$vp_nr == "VP10" & data.protocol$day == "A")], data.protocol$sampling_time_3[which(data.protocol$vp_nr == "VP10" & data.protocol$day == "A")], data.protocol$sampling_time_4[which(data.protocol$vp_nr == "VP10" & data.protocol$day == "A")])
subset.cort.long$date_D1[which(subset.cort.long$vp_nr == "VP10" & subset.cort.long$day == "A")] <- "2025-06-03"
subset.cort.long$awakening_time_D1_app[which(subset.cort.long$vp_nr == "VP10" & subset.cort.long$day == "A")] <- subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP10" & subset.cort.long$day == "A" & subset.cort.long$variable == "Cort_1")]

## VP10, day B
subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP10" & subset.cort.long$day == "B")] <- c(data.protocol$sampling_time_1[which(data.protocol$vp_nr == "VP10" & data.protocol$day == "B")], data.protocol$sampling_time_2[which(data.protocol$vp_nr == "VP10" & data.protocol$day == "B")], data.protocol$sampling_time_3[which(data.protocol$vp_nr == "VP10" & data.protocol$day == "B")], data.protocol$sampling_time_4[which(data.protocol$vp_nr == "VP10" & data.protocol$day == "B")])
subset.cort.long$date_D1[which(subset.cort.long$vp_nr == "VP10" & subset.cort.long$day == "B")] <- "2025-06-05"
subset.cort.long$awakening_time_D1_app[which(subset.cort.long$vp_nr == "VP10" & subset.cort.long$day == "B")] <- subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP10" & subset.cort.long$day == "B" & subset.cort.long$variable == "Cort_1")]

## VP10, day C
subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP10" & subset.cort.long$day == "C")] <- c(data.protocol$sampling_time_1[which(data.protocol$vp_nr == "VP10" & data.protocol$day == "C")], data.protocol$sampling_time_2[which(data.protocol$vp_nr == "VP10" & data.protocol$day == "C")], data.protocol$sampling_time_3[which(data.protocol$vp_nr == "VP10" & data.protocol$day == "C")], data.protocol$sampling_time_4[which(data.protocol$vp_nr == "VP10" & data.protocol$day == "C")])
subset.cort.long$date_D1[which(subset.cort.long$vp_nr == "VP10" & subset.cort.long$day == "C")] <- "2025-06-10"
subset.cort.long$awakening_time_D1_app[which(subset.cort.long$vp_nr == "VP10" & subset.cort.long$day == "C")] <- subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP10" & subset.cort.long$day == "C" & subset.cort.long$variable == "Cort_1")]

## VP10, day D
subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP10" & subset.cort.long$day == "D")] <- c(data.protocol$sampling_time_1[which(data.protocol$vp_nr == "VP10" & data.protocol$day == "D")], data.protocol$sampling_time_2[which(data.protocol$vp_nr == "VP10" & data.protocol$day == "D")], data.protocol$sampling_time_3[which(data.protocol$vp_nr == "VP10" & data.protocol$day == "D")], data.protocol$sampling_time_4[which(data.protocol$vp_nr == "VP10" & data.protocol$day == "D")])
subset.cort.long$date_D1[which(subset.cort.long$vp_nr == "VP10" & subset.cort.long$day == "D")] <- "2025-06-12"
subset.cort.long$awakening_time_D1_app[which(subset.cort.long$vp_nr == "VP10" & subset.cort.long$day == "D")] <- subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP10" & subset.cort.long$day == "D" & subset.cort.long$variable == "Cort_1")]

## no available log data: 11C

## VP12, day B
subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP12" & subset.cort.long$day == "B")] <- c(data.protocol$sampling_time_1[which(data.protocol$vp_nr == "VP12" & data.protocol$day == "B")], data.protocol$sampling_time_2[which(data.protocol$vp_nr == "VP12" & data.protocol$day == "B")], data.protocol$sampling_time_3[which(data.protocol$vp_nr == "VP12" & data.protocol$day == "B")], data.protocol$sampling_time_4[which(data.protocol$vp_nr == "VP12" & data.protocol$day == "B")])
subset.cort.long$date_D1[which(subset.cort.long$vp_nr == "VP12" & subset.cort.long$day == "B")] <- "2025-06-05"
subset.cort.long$awakening_time_D1_app[which(subset.cort.long$vp_nr == "VP12" & subset.cort.long$day == "B")] <- subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP12" & subset.cort.long$day == "B" & subset.cort.long$variable == "Cort_1")]

## VP12, day C
subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP12" & subset.cort.long$day == "C")] <- c(data.protocol$sampling_time_1[which(data.protocol$vp_nr == "VP12" & data.protocol$day == "C")], data.protocol$sampling_time_2[which(data.protocol$vp_nr == "VP12" & data.protocol$day == "C")], data.protocol$sampling_time_3[which(data.protocol$vp_nr == "VP12" & data.protocol$day == "C")], data.protocol$sampling_time_4[which(data.protocol$vp_nr == "VP12" & data.protocol$day == "C")])
subset.cort.long$date_D1[which(subset.cort.long$vp_nr == "VP12" & subset.cort.long$day == "C")] <- "2025-06-10"
subset.cort.long$awakening_time_D1_app[which(subset.cort.long$vp_nr == "VP12" & subset.cort.long$day == "C")] <- subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP12" & subset.cort.long$day == "C" & subset.cort.long$variable == "Cort_1")]

## VP12, day D
subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP12" & subset.cort.long$day == "D")] <- c(data.protocol$sampling_time_1[which(data.protocol$vp_nr == "VP12" & data.protocol$day == "D")], data.protocol$sampling_time_2[which(data.protocol$vp_nr == "VP12" & data.protocol$day == "D")], data.protocol$sampling_time_3[which(data.protocol$vp_nr == "VP12" & data.protocol$day == "D")], data.protocol$sampling_time_4[which(data.protocol$vp_nr == "VP12" & data.protocol$day == "D")])
subset.cort.long$date_D1[which(subset.cort.long$vp_nr == "VP12" & subset.cort.long$day == "D")] <- "2025-06-12"
subset.cort.long$awakening_time_D1_app[which(subset.cort.long$vp_nr == "VP12" & subset.cort.long$day == "D")] <- subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP12" & subset.cort.long$day == "D" & subset.cort.long$variable == "Cort_1")]

## VP15, day D
subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP15" & subset.cort.long$day == "D")] <- c(data.protocol$sampling_time_1[which(data.protocol$vp_nr == "VP15" & data.protocol$day == "D")], data.protocol$sampling_time_2[which(data.protocol$vp_nr == "VP15" & data.protocol$day == "D")], data.protocol$sampling_time_3[which(data.protocol$vp_nr == "VP15" & data.protocol$day == "D")], data.protocol$sampling_time_4[which(data.protocol$vp_nr == "VP15" & data.protocol$day == "D")])
subset.cort.long$date_D1[which(subset.cort.long$vp_nr == "VP15" & subset.cort.long$day == "D")] <- "2025-06-26"
subset.cort.long$awakening_time_D1_app[which(subset.cort.long$vp_nr == "VP15" & subset.cort.long$day == "D")] <- subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP15" & subset.cort.long$day == "D" & subset.cort.long$variable == "Cort_1")]

## VP17, day B
subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP17" & subset.cort.long$day == "B")] <- c(data.protocol$sampling_time_1[which(data.protocol$vp_nr == "VP17" & data.protocol$day == "B")], data.protocol$sampling_time_2[which(data.protocol$vp_nr == "VP17" & data.protocol$day == "B")], data.protocol$sampling_time_3[which(data.protocol$vp_nr == "VP17" & data.protocol$day == "B")], data.protocol$sampling_time_4[which(data.protocol$vp_nr == "VP17" & data.protocol$day == "B")])
subset.cort.long$date_D1[which(subset.cort.long$vp_nr == "VP17" & subset.cort.long$day == "B")] <- "2025-06-19"
subset.cort.long$awakening_time_D1_app[which(subset.cort.long$vp_nr == "VP17" & subset.cort.long$day == "B")] <- subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP17" & subset.cort.long$day == "B" & subset.cort.long$variable == "Cort_1")]

## VP24, day A (only awakening is wrong)
subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP24" & subset.cort.long$day == "A" & subset.cort.long$variable == "Cort_1")] <- data.protocol$sampling_time_1[which(data.protocol$vp_nr == "VP24" & data.protocol$day == "A")] 
subset.cort.long$awakening_time_D1_app[which(subset.cort.long$vp_nr == "VP24" & subset.cort.long$day == "A")] <- subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP24" & subset.cort.long$day == "A" & subset.cort.long$variable == "Cort_1")]

## VP35, day "C"
subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP35" & subset.cort.long$day == "C")] <- c(data.protocol$sampling_time_1[which(data.protocol$vp_nr == "VP35" & data.protocol$day == "C")], data.protocol$sampling_time_2[which(data.protocol$vp_nr == "VP35" & data.protocol$day == "C")], data.protocol$sampling_time_3[which(data.protocol$vp_nr == "VP35" & data.protocol$day == "C")], data.protocol$sampling_time_4[which(data.protocol$vp_nr == "VP35" & data.protocol$day == "C")])
subset.cort.long$date_D1[which(subset.cort.long$vp_nr == "VP35" & subset.cort.long$day == "C")] <- "2025-07-29"
subset.cort.long$awakening_time_D1_app[which(subset.cort.long$vp_nr == "VP35" & subset.cort.long$day == "C")] <- subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP35" & subset.cort.long$day == "C" & subset.cort.long$variable == "Cort_1")]

## VP35, day "D"
subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP35" & subset.cort.long$day == "D")] <- c(data.protocol$sampling_time_1[which(data.protocol$vp_nr == "VP35" & data.protocol$day == "D")], data.protocol$sampling_time_2[which(data.protocol$vp_nr == "VP35" & data.protocol$day == "D")], data.protocol$sampling_time_3[which(data.protocol$vp_nr == "VP35" & data.protocol$day == "D")], data.protocol$sampling_time_4[which(data.protocol$vp_nr == "VP35" & data.protocol$day == "D")])
subset.cort.long$date_D1[which(subset.cort.long$vp_nr == "VP35" & subset.cort.long$day == "D")] <- "2025-07-31"
subset.cort.long$awakening_time_D1_app[which(subset.cort.long$vp_nr == "VP35" & subset.cort.long$day == "D")] <- subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP35" & subset.cort.long$day == "D" & subset.cort.long$variable == "Cort_1")]

## VP38, day "B"
subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP38" & subset.cort.long$day == "B")] <- c(data.protocol$sampling_time_1[which(data.protocol$vp_nr == "VP38" & data.protocol$day == "B")], data.protocol$sampling_time_2[which(data.protocol$vp_nr == "VP38" & data.protocol$day == "B")], data.protocol$sampling_time_3[which(data.protocol$vp_nr == "VP38" & data.protocol$day == "B")], data.protocol$sampling_time_4[which(data.protocol$vp_nr == "VP38" & data.protocol$day == "B")])
subset.cort.long$date_D1[which(subset.cort.long$vp_nr == "VP38" & subset.cort.long$day == "B")] <- "2025-08-07"
subset.cort.long$awakening_time_D1_app[which(subset.cort.long$vp_nr == "VP38" & subset.cort.long$day == "B")] <- subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP38" & subset.cort.long$day == "B" & subset.cort.long$variable == "Cort_1")]

## "VP47", day "A"
subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP47" & subset.cort.long$day == "A" & subset.cort.long$variable == "Cort_2")] <- data.protocol$sampling_time_2[which(data.protocol$vp_nr == "VP47" & data.protocol$day == "A")]
subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP47" & subset.cort.long$day == "A" & subset.cort.long$variable == "Cort_3")] <- data.protocol$sampling_time_3[which(data.protocol$vp_nr == "VP47" & data.protocol$day == "A")]
subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP47" & subset.cort.long$day == "A" & subset.cort.long$variable == "Cort_4")] <- data.protocol$sampling_time_4[which(data.protocol$vp_nr == "VP47" & data.protocol$day == "A")]
subset.cort.long$date_D1[which(subset.cort.long$vp_nr == "VP47" & subset.cort.long$day == "A")] <- "2025-08-19"
subset.cort.long$awakening_time_D1_app[which(subset.cort.long$vp_nr == "VP47" & subset.cort.long$day == "A")] <- paste(subset.cort.long$date_D1[which(subset.cort.long$vp_nr == "VP47" & subset.cort.long$day == "A")], "06:43:11", sep = " ") 

## VP50, day "B"
subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP50" & subset.cort.long$day == "B")] <- c(data.protocol$sampling_time_1[which(data.protocol$vp_nr == "VP50" & data.protocol$day == "B")], data.protocol$sampling_time_2[which(data.protocol$vp_nr == "VP50" & data.protocol$day == "B")], data.protocol$sampling_time_3[which(data.protocol$vp_nr == "VP50" & data.protocol$day == "B")], data.protocol$sampling_time_4[which(data.protocol$vp_nr == "VP50" & data.protocol$day == "B")])
subset.cort.long$date_D1[which(subset.cort.long$vp_nr == "VP50" & subset.cort.long$day == "B")] <- "2025-08-29"
subset.cort.long$awakening_time_D1_app[which(subset.cort.long$vp_nr == "VP50" & subset.cort.long$day == "B")] <- subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP50" & subset.cort.long$day == "B" & subset.cort.long$variable == "Cort_1")]

## VP50, day "C"
subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP50" & subset.cort.long$day == "C")] <- c(data.protocol$sampling_time_1[which(data.protocol$vp_nr == "VP50" & data.protocol$day == "C")], data.protocol$sampling_time_2[which(data.protocol$vp_nr == "VP50" & data.protocol$day == "C")], data.protocol$sampling_time_3[which(data.protocol$vp_nr == "VP50" & data.protocol$day == "C")], data.protocol$sampling_time_4[which(data.protocol$vp_nr == "VP50" & data.protocol$day == "C")])
subset.cort.long$date_D1[which(subset.cort.long$vp_nr == "VP50" & subset.cort.long$day == "C")] <- "2025-09-02"
subset.cort.long$awakening_time_D1_app[which(subset.cort.long$vp_nr == "VP50" & subset.cort.long$day == "C")] <- subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP50" & subset.cort.long$day == "C" & subset.cort.long$variable == "Cort_1")]

## VP50, day "D"
subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP50" & subset.cort.long$day == "D")] <- c(data.protocol$sampling_time_1[which(data.protocol$vp_nr == "VP50" & data.protocol$day == "D")], data.protocol$sampling_time_2[which(data.protocol$vp_nr == "VP50" & data.protocol$day == "D")], data.protocol$sampling_time_3[which(data.protocol$vp_nr == "VP50" & data.protocol$day == "D")], data.protocol$sampling_time_4[which(data.protocol$vp_nr == "VP50" & data.protocol$day == "D")])
subset.cort.long$date_D1[which(subset.cort.long$vp_nr == "VP50" & subset.cort.long$day == "D")] <- "2025-09-04"
subset.cort.long$awakening_time_D1_app[which(subset.cort.long$vp_nr == "VP50" & subset.cort.long$day == "D")] <- subset.cort.long$sampling_time[which(subset.cort.long$vp_nr == "VP50" & subset.cort.long$day == "D" & subset.cort.long$variable == "Cort_1")]

## VP 4 day C Cort 1 and 4
# Cort 1: correct time = 07:51:59	
subset.cort.long$sampling_time[subset.cort.long$vp_nr == "VP04" & subset.cort.long$day == "C" & subset.cort.long$variable == "Cort_1"] <- paste(subset.cort.long$date_D1[subset.cort.long$vp_nr == "VP04" & subset.cort.long$day == "C" & subset.cort.long$variable == "Cort_1"], " 07:51:59")
# Cort 4: correct time = 08:51:20
subset.cort.long$sampling_time[subset.cort.long$vp_nr == "VP04" & subset.cort.long$day == "C" & subset.cort.long$variable == "Cort_4"] <- paste(subset.cort.long$date_D1[subset.cort.long$vp_nr == "VP04" & subset.cort.long$day == "C" & subset.cort.long$variable == "Cort_1"], " 08:51:20") 

## VP 12 day A Cort 1 and 4
# Cort 1: correct time = 06:32:28
subset.cort.long$sampling_time[subset.cort.long$vp_nr == "VP12" & subset.cort.long$day == "A" & subset.cort.long$variable == "Cort_1"] <- paste(subset.cort.long$date_D1[subset.cort.long$vp_nr == "VP12" & subset.cort.long$day == "A" & subset.cort.long$variable == "Cort_1"], " 06:32:28") 

# Cort 4: correct time = 07:32:27
subset.cort.long$sampling_time[subset.cort.long$vp_nr == "VP12" & subset.cort.long$day == "A" & subset.cort.long$variable == "Cort_4"] <- paste(subset.cort.long$date_D1[subset.cort.long$vp_nr == "VP12" & subset.cort.long$day == "A" & subset.cort.long$variable == "Cort_4"], " 07:32:27")

## VP 22 day A Cort 2 and 4
# Cort 2: correct time = 07:01:55
subset.cort.long$sampling_time[subset.cort.long$vp_nr == "VP22" & subset.cort.long$day == "A" & subset.cort.long$variable == "Cort_2"] <- paste(subset.cort.long$date_D1[subset.cort.long$vp_nr == "VP22" & subset.cort.long$day == "A" & subset.cort.long$variable == "Cort_2"], " 07:01:55")
# Cort 4: correct time = 07:28:09
subset.cort.long$sampling_time[subset.cort.long$vp_nr == "VP22" & subset.cort.long$day == "A" & subset.cort.long$variable == "Cort_4"] <- paste(subset.cort.long$date_D1[subset.cort.long$vp_nr == "VP22" & subset.cort.long$day == "A" & subset.cort.long$variable == "Cort_2"], " 07:28:09")

## VP 26 day C Cort 2 and 4
# Cort 3: correct time = 08:45:52
subset.cort.long$sampling_time[subset.cort.long$vp_nr == "VP26" & subset.cort.long$day == "C" & subset.cort.long$variable == "Cort_3"] <- paste(subset.cort.long$date_D1[subset.cort.long$vp_nr == "VP26" & subset.cort.long$day == "C" & subset.cort.long$variable == "Cort_3"], " 08:45:52")
# Cort 4: correct time = 09:02:19
subset.cort.long$sampling_time[subset.cort.long$vp_nr == "VP26" & subset.cort.long$day == "C" & subset.cort.long$variable == "Cort_4"] <- paste(subset.cort.long$date_D1[subset.cort.long$vp_nr == "VP26" & subset.cort.long$day == "C" & subset.cort.long$variable == "Cort_4"], " 09:02:19")

## VP 36 day A Cort 2 and 3
# Cort 2: correct time = 10:18:29	
subset.cort.long$sampling_time[subset.cort.long$vp_nr == "VP36" & subset.cort.long$day == "A" & subset.cort.long$variable == "Cort_2"] <- paste(subset.cort.long$date_D1[subset.cort.long$vp_nr == "VP36" & subset.cort.long$day == "A" & subset.cort.long$variable == "Cort_2"], " 10:18:29")
# Cort 3: correct time = 10:33:16	
subset.cort.long$sampling_time[subset.cort.long$vp_nr == "VP36" & subset.cort.long$day == "A" & subset.cort.long$variable == "Cort_3"] <- paste(subset.cort.long$date_D1[subset.cort.long$vp_nr == "VP36" & subset.cort.long$day == "A" & subset.cort.long$variable == "Cort_2"], " 10:33:16")

## VP 14 day B Cort 3: is sampled "1 min before Cort 4" but is way higher, no way there was just one min in between
# Assume that alarm was disabled too late (9:15:00 instead of 09:30:56)
subset.cort.long$sampling_time[subset.cort.long$vp_nr == "VP14" & subset.cort.long$day == "B" & subset.cort.long$variable == "Cort_3"] <- "2025-06-19 09:15:00"

## VP18C - alarm went off in the middle of the night - awakening time D1 app wrong; assume: 2026-03-06 08:12:06
subset.cort.long$awakening_time_D1_app[subset.cort.long$vp_nr == "VP18" & subset.cort.long$day == "C"] <- "2025-07-02 08:12:06"

## VP22B - alarm went off in the middle of the night
subset.cort.long$awakening_time_D1_app[subset.cort.long$vp_nr == "VP22" & subset.cort.long$day == "B"] <- NA
subset.cort.long$sampling_time[subset.cort.long$vp_nr == "VP22" & subset.cort.long$day == "B"] <- NA

## VP39B - Cort_0 is produced - which should read Cort_4
subset.cort.long$date_D1[subset.cort.long$vp_nr == "VP39" & subset.cort.long$day == "B" & subset.cort.long$variable == "Cort_4"] <- "2025-08-07"
subset.cort.long$awakening_time_D1_app[subset.cort.long$vp_nr == "VP39" & subset.cort.long$day == "B" & subset.cort.long$variable == "Cort_4"] <- "2025-08-07 05:46:51"
subset.cort.long$sampling_time[subset.cort.long$vp_nr == "VP39" & subset.cort.long$day == "B" & subset.cort.long$variable == "Cort_4"] <- "2025-08-07 06:47:07"

subset.cort.long <- subset.cort.long[!(subset.cort.long$vp_nr == "VP39" & subset.cort.long$day == "B" & subset.cort.long$variable == "Cort_0"),]

subset.cort.long$sampling_time <- as.POSIXct(subset.cort.long$sampling_time)

# correct lag 
subset.cort.long$lag <- as.numeric((subset.cort.long$sampling_time - subset.cort.long$awakening_time_D1_app)/60)
subset.cort.long$lag_min <- subset.cort.long$lag - subset.cort.long$opt_lag

rm(logs)

####### exclude cortisol values non compliant participants
subset.cort.long$sampling_time[subset.cort.long$vp_nr == "VP34" & 
                             subset.cort.long$day == "A"] <- NA # not compliant
subset.cort.long$cortisol[subset.cort.long$vp_nr == "VP50" & 
                             subset.cort.long$day == "C"] <- NA # 2 missings

subset.cort.long$cortisol[subset.cort.long$vp_nr == "VP23" &
                             subset.cort.long$day == "D"] <- NA # extremely high awakening sample: 49.9422

subset.cort.long$sampling_time[subset.cort.long$vp_nr == "VP41" & subset.cort.long$day == "A"] <- c(NA, NA, "2025-08-12 06:15:24 CEST", "2025-08-12 06:31:54 CEST") # time logging unclear

# exclusion of those with problems (stevia)
protocol_raw <- read_excel("/Users/maria/Nextcloud/AG Pruessner/Projekte/SuCAR/5. Data and analysis/1 Rawdata/Tagebuch/SuCAR_Messtagebuch_20251007.xls")
protocol_raw <- protocol_raw[-1, ]
protocol_raw$subject_number <- as.factor(as.numeric(substr(protocol_raw$number, start = 3, stop = 4)))

protocol_raw$number[c(6, 7, 34, 46)]
protocol_raw$day[c(6, 7, 34, 46)]

subset.cort.long$cortisol[subset.cort.long$subject_number == "VP02" & 
                            subset.cort.long$day == "B"] <- NA # allergische Reaktion
subset.cort.long$cortisol[subset.cort.long$subject_number == "VP02" & 
                            subset.cort.long$day == "C"] <- NA # allergische Reaktion
subset.cort.long$cortisol[subset.cort.long$subject_number == "VP10" & 
                            subset.cort.long$day == "A"] <- NA # allergische Reaktion
subset.cort.long$cortisol[subset.cort.long$subject_number == "VP13" & 
                            subset.cort.long$day == "A"] <- NA # allergische Reaktion

# remove the NAs in Cort
subset.cort.long <- subset.cort.long[!is.na(subset.cort.long$cortisol), ]

# 21 A has two Cort_1 entries - remove first entry
subset.cort.long <- subset.cort.long[-305, ]

# check if each participant has 4 measurements per sampling day
table(subset.cort.long$vp_nr, subset.cort.long$day)

# compute AUCi
source("~/Documents/R_extensions/R_auc03_20220410.R")
subset.cort.long$AUCi_cort <- NA
# for (ix in 1:length(subset.cort.long$lag)) {
#   if(is.na(subset.cort.long$lag[ix])) {
#     subset.cort.long$lag[ix] <- subset.cort.long$opt_lag[ix]
#   } else {
#     subset.cort.long$lag[ix] <- subset.cort.long$lag[ix]
#   }
# }

# compute AUCi with opt_lag 
for (i in unique(subset.cort.long$vp_nr)) {
  current.set <- subset.cort.long[subset.cort.long$vp_nr == i,]
  for (indx in unique(current.set$day)) {
  current.day <- current.set[current.set$day == indx,] 
  if(sum(is.na(current.set$cortisol)) > 0) {
    subset.cort.long$AUCi_cort[subset.cort.long$vp_nr == i & subset.cort.long$day == indx] <- NA
  } else {
    subset.cort.long$AUCi_cort[subset.cort.long$vp_nr == i & subset.cort.long$day == indx] <- auc03(current.day$opt_lag, current.day$cortisol, type = "i")
  } 
  }
}
rm(current.day); rm(current.set)

# write csv
write.csv(x = subset.cort.long, 
          file = paste(project, "_cort_long_", Sys.Date(), ".csv", sep = ""))

subset.cort.long$condition <- as.factor(subset.cort.long$condition)

subset.cort.wide <- data.frame(matrix(data = NA, nrow = length(unique(subset.cort.long$vp_nr)), ncol = 3))
names(subset.cort.wide) <- c("vp_nr", "AUCi_cort_stevia", "AUCi_cort_glucose")
subset.cort.wide$vp_nr <- unique(subset.cort.long$vp_nr)
for (i in unique(subset.cort.long$vp_nr)) {
  subset.cort.wide$AUCi_cort_stevia[subset.cort.wide$vp_nr == i] <- mean(subset.cort.long$AUCi_cort[subset.cort.long$vp_nr == i &
                                                                         subset.cort.long$condition == "stevia"], na.rm = TRUE)

  subset.cort.wide$AUCi_cort_glucose[subset.cort.wide$vp_nr == i] <- mean(subset.cort.long$AUCi_cort[subset.cort.long$vp_nr == i &
                                                                         subset.cort.long$condition == "glucose"], na.rm = TRUE)
}

test <- merge(subset.cort.wide, unique(data.protocol[, c("vp_nr", "sugar_amount")]), by = c("vp_nr"))

# Erstellen der ABBILDUNGEN

# PLOT Hypothese 1
subset.cort.long$day <- factor(subset.cort.long$day)

# Calculate mean and standard error for each day
summary_data <- subset.cort.long %>%
  group_by(condition) %>%
  dplyr::summarize(
    mean_Cort_AUC = mean(AUCi_cort),
    se = sd(AUCi_cort) / sqrt(length(AUCi_cort))) %>%
  mutate(ymin = mean_Cort_AUC - se,
         ymax = mean_Cort_AUC + se)

# Create the bar plot with error bars
A <- ggplot(summary_data, aes(x = condition, y = mean_Cort_AUC)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7, 
           fill = "skyblue") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), 
                position = position_dodge(width = 0.7), 
                width = 0.2) +
  labs(title = "A",
       x = "day",
       y = "Cortisol Awakening Response\n(area under the curve)") +
  theme_classic()
A
```


### CARWatch Log Files

### Libre 3 Glucose


### Cortisol Sweet Memories

```{r}
# create subdirectory for cortisol in 2 Cleaned Data
dir.create("/Users/maria/Nextcloud/AG Pruessner/Projekte/Sweet Memories I/5. Data and analysis/2 Cleaned Data/Cortisol/")
setwd("/Users/maria/Nextcloud/AG Pruessner/Projekte/Sweet Memories I/5. Data and analysis/2 Cleaned Data/Cortisol/")

# read in excel you received from wetlab
cortisol_raw <- readxl::read_excel("/Users/maria/Nextcloud/AG Pruessner/Projekte/Sweet Memories I/5. Data and analysis/1 Rawdata/Cortisol/SweetMemories_cort_20240214.xlsm")

# read out inter and intra CV
interCV <- round(mean(as.numeric(names(cortisol_raw)[6])), 2)

intraCV_col <- cortisol_raw[, 8]
names(intraCV_col) <- "intra"
intraCV_col$intra[intraCV_col$intra == "intra %CV"] <- NA
intraCV_col$intra <- as.numeric(intraCV_col$intra)
intraCV <- round(mean(intraCV_col$intra, na.rm = TRUE), 2)
rm(intraCV_col)

CV_data <- data.frame(matrix(data = NA, nrow = 1, ncol = 2)); names(CV_data) <- c("interCV", "intraCV")
CV_data$interCV <- interCV; CV_data$intraCV <- intraCV 

# save inter and intra CV means in file
write.csv(x = CV_data, 
          file = paste("/Users/maria/Nextcloud/AG Pruessner/Projekte/Sweet Memories I/5. Data and analysis/2 Cleaned Data/Cortisol/SweetMemories_cort_assay_CVs_", Sys.Date(), ".csv", sep = ""))

# write cortisol data into dataframe (long) - csv and RData
subset.cort.long <- cortisol_raw[, c(2, 3, 7)]

colnames(subset.cort.long) <- c("vp_number", "variable", "cortisol")
subset.cort.long <- subset.cort.long[-which(subset.cort.long$vp_number %in% c("Label", NA)),]

#subset.cort.long$code <- toupper(subset.cort.long$code)
#subset.cort.long$code[subset.cort.long$code == "CI21"] <- "CI12"
#subset.cort.long$code[subset.cort.long$code == "VI15"] <- "UI15"

subset.cort.long$vp_number <- parse_number(subset.cort.long$vp_number)
subset.cort.long$vp_number[subset.cort.long$vp_number == "143"] <- "143/62"
subset.cort.long$vp_number[subset.cort.long$vp_number == "61"] <- "63"

#subset.cort.long$code <- as.factor(subset.cort.long$code)
subset.cort.long$cortisol <- as.numeric(subset.cort.long$cortisol)
subset.cort.long$variable <- paste("Cort_", subset.cort.long$variable, sep = "")

write.csv(x = subset.cort.long, 
          file = paste("/Users/maria/Nextcloud/AG Pruessner/Projekte/Sweet Memories I/5. Data and analysis/2 Cleaned Data/Cortisol/SweetMemories_cort_long_", Sys.Date(), ".csv", sep = ""))

# write cortisol data into dataframe (wide) - csv and RData
library(reshape2)
data.cortisol <- dcast(data = subset.cort.long, vp_number ~ variable, value.var = "cortisol")

write.csv(x = data.cortisol, 
          file = paste("/Users/maria/Nextcloud/AG Pruessner/Projekte/Sweet Memories I/5. Data and analysis/2 Cleaned Data/Cortisol/SweetMemories_cort_wide_", Sys.Date(), ".csv", sep = ""))

rm(cortisol_raw); rm(CV_data)
```



# Processing of the variables needed for the analysis

## Continuous glucose recordings

**Glucose levels.** A tissue glucose measurement is carried out continuously over 14 days with the help of a FreeStyle Libre 3 sensor (FreeStyle Libre, Berlin, Germany) and the Libre 3 application. Tissue glucose correlates with the blood glucose level and allows a minute-wise measurement without repeated lancing. 

## Salivary cortisol

Cortisol Awakening Response (CAR). The CAR is assessed on four days (when the drinks are consumed) by using four salivary samples that are collected upon waking up (prior to drink consumption) and +30, +45 and +60 minutes later. Compliance of the CAR assessment will be checked using the CARwatch application. For statistical analyses, we will calculate the mean across each timepoint (awakening, +30, +45 and +60 minutes) per drink condition (glucose, sweetener) to enhance the reliability of the assessment. 

## Sex and menstrual cycle phase

Sex (assigned at birth) and menstrual cycle phase. These variables are assessed via self-report. For females, menstrual cycle phase (follicular phase, luteal phase) will be estimated following current recommendations 4 based on the self-reported day of the last cycle start and average cycle length as well as an estimated 14 days period for the luteal phase. 

## Potential covariates

Participants fill in a measurement diary on each CAR assessment day to record environmental and personal state factors during the assessment (e.g., light conditions during awakening, use of alarm clock, health related variables, physical activity) 2. We will check whether these variables are associated with the respective outcome of our analysis and if it is the case, enter them as covariates in the statistical models. 

Apart from these measures, participants fill in different questionnaires (e.g., on chronic stress, food craving, physical activity), which are not part of the current preregistration, but used to describe the sample. 

### Labor T1

```{r load lab day 1 data questionstar}
# Lab day 1: code
data.quest1 <- read.csv("/Users/maria/Nextcloud/AG Pruessner/Projekte/SuCAR/5. Data and analysis/1 Rawdata/Questionstar/Labor T1/SuCAR Laborfragebogen - alle-Antworten 2025-05-09_numbers.csv", na.strings = "NULL")

data.quest1 <- data.quest1[-which(is.na(data.quest1$code)), -c(1:8, 10:50, 333:338)]
names(data.quest1)[names(data.quest1) == "S14"] <- "date_lab1"
names(data.quest1) <- gsub("\\.", "_", names(data.quest1))

# extract date lab 1
# data.quest1$date_lab1 <- strptime(x = data.quest1$date_lab1,
#                                     format = "%dd.%mm.$yy %hh:%mm", 
#                                     tz = "CET")

data.quest1$code <- toupper(data.quest1$code)
data.quest1$code_2 <- NULL

# correct wrong vp_nrs and codes
# VP50 code BA08 is actually VP49 - change this 
data.quest1$vp_nr[data.quest1$code == "BA08"] <- "VP49"

# ASK DANA WHY SHE DID THIS?
# # adjust duplicate VP22: code VN05 -> VP51 (completed just the first laboratory session)
# TICS.filtered$vp_nr[TICS.filtered$code == "VN05"] <- "VP51"
```

Education
```{r}
# Education (years)
data.quest1$education_2 <- as.numeric(str_replace(data.quest1$education_2, 
                                pattern = ",", replacement = "."))
names(data.quest1)[names(data.quest1) == "education_2"] <- "education_yrs"
```

BMI
```{r BMI}
data.quest1$BMI <- (as.numeric(data.quest1$weight_1)/as.numeric(data.quest1$height_1)^2)*10000
```

Compliance
```{r compliance lab 1}
data.quest1$coffee[data.quest1$coffee == "1-2"] <- 1.5
data.quest1$coffee[data.quest1$coffee %in% c("2-3", "2–3")] <- 2.5
data.quest1$coffee[data.quest1$coffee %in% c("0-1", "1/2", "0,5")] <- 0.5
data.quest1$coffee[data.quest1$coffee %in% c("0–2", "0-2")] <- 1
data.quest1$coffee <- as.numeric(data.quest1$coffee)

data.quest1$exercise[data.quest1$exercise %in% c("3,5", "3-4")] <- 3.5
data.quest1$exercise[data.quest1$exercise %in% c("3h", "2–4")] <- 3
data.quest1$exercise[data.quest1$exercise %in% c("5-6")] <- 5.5
data.quest1$exercise[data.quest1$exercise %in% c("1,5")] <- 1.5
data.quest1$exercise[data.quest1$exercise %in% c("4-5")] <- 4.5
data.quest1$exercise <- as.numeric(data.quest1$exercise)

data.quest1$lab_state_disease_1_O <- NULL
```

Menstrual cycle phase (duration)
```{r menstrual cycle}
# tidy duration variable
data.quest1$lab_state_cycle_dur[data.quest1$lab_state_cycle_dur == "26-28"] <- 27
data.quest1$lab_state_cycle_dur <- as.numeric(data.quest1$lab_state_cycle_dur)
```

```{r}
test <- merge(test, data.quest1[, c("sex", "vp_nr", "code", "age")], by = "vp_nr")

test2 <- reshape::melt(test, 
                       measure.vars = c("AUCi_cort_stevia",
                                        "AUCi_cort_glucose"))
test2$sex <- as.factor(test2$sex)

test3 <- test2[test2$sugar_amount == 50,]
randomIntercept <- lme(scale(value) ~ variable * sex, 
                       data = test2, random = ~variable|code, 
                       method = "ML", 
                       na.action = na.omit, 
                       control = list(opt = "optim"))
sjPlot::tab_model(randomIntercept, show.se = TRUE, show.ci = FALSE)
```


#### Kurzform des Trierer Inventars zur Messung von chronischem Stress (TICS-9)

```{r TICS-sf}
#Likert scale 1:5
#missing value imputation: median
#if more than 3 (of 9) values are missing *for sum score*: NA
#Higher scores suggest higher chronic stress

TICS.raw <- data.quest1[ , c("code", "TICS_sf_1", "TICS_sf_2", "TICS_sf_3", "TICS_sf_4", "TICS_sf_5","TICS_sf_6", "TICS_sf_7", "TICS_sf_8", "TICS_sf_9")] 

TICS.raw[2:10] <- sapply(TICS.raw[2:10], as.numeric)

# Correcting invalid values 
str(TICS.raw) # inspect TICS raw data for  range of data (always 1 - 5)

which((TICS.raw[ , c("TICS_sf_1", "TICS_sf_2", "TICS_sf_3", "TICS_sf_4", "TICS_sf_5", "TICS_sf_6", "TICS_sf_7", "TICS_sf_8", "TICS_sf_9")] < 1) | (TICS.raw[ , c("TICS_sf_1", "TICS_sf_2", "TICS_sf_3", "TICS_sf_4", "TICS_sf_5", "TICS_sf_6", "TICS_sf_7", "TICS_sf_8", "TICS_sf_9")] > 5)) # inspect no. of invalid item scores, if output = integer(0) --> no invalid scores, ignore next step

# TICS.raw[TICS.raw < 1] <- NA # first check data entry for mistakes!! 
# TICS.raw[TICS.raw > 5] <- NA # otherwise recode invalid values as missings (NA)

table(is.na(TICS.raw)) # inspect no. of NA-items

# Calculation of Sum Score 
key.list.TICS <- list(TICS_sum = c("TICS_sf_1", "TICS_sf_2", "TICS_sf_3", "TICS_sf_4", "TICS_sf_5","TICS_sf_6", "TICS_sf_7", "TICS_sf_8", "TICS_sf_9")) 
keys.TICS <- make.keys(TICS.raw, key.list.TICS)

TICS.score.res <- scoreItems(keys.TICS, TICS.raw, totals = TRUE, missing = TRUE, impute = "median", delete = FALSE)
# for scoreItems: change totals = FALSE for average scores (default); change impute for different NA-replacement )"mean", "median", "none")

TICS.raw <- cbind(TICS.raw, TICS.score.res$scores)

# add a loop to switch missing cases to NA
# sum score
for(i in 1:nrow(TICS.raw)){
  if (length(which(is.na(TICS.raw[i,]))) > 3) {
    TICS.raw$TICS_sum[i] <- NA
  }
}
rm(i)

# Gütekriterien
TICS_stats <- data.frame(
  scale = "TICS_sum",
  alpha = round(TICS.score.res$alpha, 2)
)

# Anzeige in der Konsole
print(TICS_stats)

# Exportiere als CSV
write.csv(TICS_stats, paste0(wd, "3 Scored Data/Questionnaires/SweetStress_TICS_stats.csv"), row.names = FALSE)

write.csv(TICS.raw, "~/Nextcloud/AG Pruessner/Projekte/Sweet Stress/5. Data and analysis/3 Scored Data/Questionnaires/SweetStress_TICS_scored.csv", row.names = FALSE)

data.quest1 <- merge(data.quest1, TICS.raw[, c("code", "TICS_sum")], by = "code")

rm(key.list.TICS); rm(keys.TICS); rm(TICS_stats); rm(TICS.raw); rm(TICS.score.res)
```

#### Fragebogen to be processed....

UCLA, RSES, ERQ, PBI, BDI, STAI-T, HEQ, BIS_BAS, IPAQ, BSA, FCQ-T, EEQ 

### Labor T2

```{r load lab day 2 data questionstar}
# Lab day 2: code
data.quest2 <- read.csv("/Users/maria/Nextcloud/AG Pruessner/Projekte/SuCAR/5. Data and analysis/1 Rawdata/Questionstar/Labor T2/SuCAR Abschluss Laborfragebogen - alle-Antworten 2025-05-09_numbers.csv", na.strings = "NULL", sep = ";")

data.quest2 <- data.quest2[, -c(1:8, 10:32, 187:192)]
names(data.quest2)[names(data.quest2) == "S14"] <- "date_lab2"
names(data.quest2)[names(data.quest2) == "Q1"] <- "vp_nr"
names(data.quest2) <- gsub("\\.", "_", names(data.quest2))
names(data.quest2) <- gsub("Q41", "BSI", names(data.quest2))
names(data.quest2) <- gsub("Q42", "DEBQ", names(data.quest2))

# extract date lab 1
# data.quest1$date_lab1 <- strptime(x = data.quest1$date_lab1,
#                                     format = "%dd.%mm.$yy %hh:%mm", 
#                                     tz = "CET")

data.quest2$code <- toupper(data.quest2$code)
data.quest2$code_2 <- toupper(data.quest2$code_2)
data.quest2$code_2 <- NULL

# correct vp_nr column
data.quest2$vp_nr[data.quest2$code == "MU02"] <- data.quest1$vp_nr[data.quest1$code == "MU02"]
data.quest2$vp_nr[data.quest2$code == "BH05"] <- data.quest1$vp_nr[data.quest1$code == "BH05"]
data.quest2$vp_nr[data.quest2$code == "BL04"] <- data.quest1$vp_nr[data.quest1$code == "BL04"]
data.quest2$vp_nr[data.quest2$code == "CL01"] <- data.quest1$vp_nr[data.quest1$code == "CL01"]
data.quest2$vp_nr[data.quest2$code == "OA27"] <- data.quest1$vp_nr[data.quest1$code == "OA27"]
data.quest2$vp_nr[data.quest2$vp_nr %in% c("10", "11" ,"12", "48")] <- paste0("VP", data.quest2$vp_nr[data.quest2$vp_nr %in% c("10", "11" ,"12", "48")])

data.quest2$code[data.quest2$vp_nr == "VP13"] <- data.quest1$code[data.quest1$vp_nr == "VP13"]
```

#### Questionnaires to be processed: PSQI, DMEQ, CTQ, IES, BSI, DEBQ etc.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Version control

```{r}
sessionInfo()
```
